---
title: "Assignment"
author: "Beltran Aller Lopez"
date: "1/11/2019"
output: 
  html_document:
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: false
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# LIBRARIES & SEED

```{r}
library("gradDescent")
library(ggplot2)
set.seed(123)
```

# FUNCTIONS

In this chapter I write the functions that I will use to solve the exercises.

## SIGMOID

```{r SIGMOID}
Sigmoid <- function(x) { 
  1 / (1 + exp(-x))
}

# feed with data
x <- seq(-5, 5, 0.01)

# and plot
plot(x, Sigmoid(x), col = 'blue', ylim = c(-.2, 1))
abline(h = 0, v = 0, col = "gray60")
```

## COST FUNCTION

```{r COST FUNCTION}
# Ref: https://www.r-bloggers.com/logistic-regression-with-r-step-by-step-implementation-part-2/
# Cost Function
# 
CostFunction <- function(parameters, X, Y) {
  n <- nrow(X)
  # function to apply (%*% Matrix multiplication)
  g <- Sigmoid(X %*% parameters)
  J <- (1/n) * sum((-Y * log(g)) - ((1 - Y) * log(1 - g)))
  return(J)
}
```

## GRADIENT DESCENT FUNCTION

```{r GRADIENT DESCENT FUNCTION}
# We want to minimize the cost function. Then derivate this funcion
TestGradientDescent <- function(iterations = 1200, X, Y) {
  
  # Initialize (b, W)
  parameters <- rep(0, ncol(X))
  # Check evolution
  print(paste("Initial Cost Function value: ", 
              convergence <- c(CostFunction(parameters, X, Y)), sep = ""))
  
  # updating (b, W) using gradient update
  
  # Derive theta using gradient descent using optim function
  # Look for information about the "optim" function (there are other options)
  parameters_optimization <- optim(par = parameters, fn = CostFunction, X = X, Y = Y, 
                                   control = list(maxit = iterations))
  #set parameters
  parameters <- parameters_optimization$par
  
  # Check evolution
  print(paste("Final Cost Function value: ", 
              convergence <- c(CostFunction(parameters, X, Y)), sep = ""))

 return(parameters) 
}
```

# DATA

I'm going tu use the next dataset which includes the scores of two examns made by a group of students, the third column indicates if they were admitted at the university. 1 means YES, 0 means NO.

```{r LOADING AND READING DATA}
#Load data
data <- read.csv("data/4_1_data.csv")
```

# EXERCISE 1

Now I create the dataframe of the predictor variables, which will be the scores of both examns and b.

```{r PREDICTOR AND RESPONSE VARIABLES}
#Predictor variables
X <- as.matrix(data[, c(1,2)])

#Add ones to X in the first column (matrix multiplication x b)
X <- cbind(rep(1, nrow(X)), X)

#Response variable
Y <- as.matrix(data$label)
```

The next step will be calculate the parameters using the TestGradientDescent function.

```{r PARAMETERS}
parameters <- TestGradientDescent(X = X, Y = Y)
parameters
```

Now I can predict the possibility of each student of being accepted in the university.

```{r PROBABILITIES OF ALL STUDENTS}
probabilities_of_all_students <- t(Sigmoid(parameters %*% t(X)))
```

Having both dataframes, the real results and the predicted results, we can do the confussion matrix and calculate the accuracy rate. In this exercise I put the cut off of the predicted results at 0.85, the accuracy is going to depend on the cut off assigned.

```{r CONFUSSION MATRIX}
confussion_matrix <- table(data$label, probabilities_of_all_students > 0.85,
                    dnn = c("Actual", "Predicted"))
confussion_matrix
```

```{r ACCURACY}
print("Accuracy:")
print(accuracy <- 1 - (confussion_matrix[2,1] + confussion_matrix[1,2])/sum(confussion_matrix))
```

# EXERCISE 2

```{r LOOP, results='hide'}

# At first I create a dataframe in which I will keep the number of iterations and the value of the          # CostFunction
iterations <- data.frame("Iterations" = 0, "CostFunctionValue" = 0) 

# Now I do a loop
for (i in seq(1:125)) {
  
  # Here we have the values of the parameters which are calculated using the TestGradientDescent function
  parameters <- TestGradientDescent(iterations = i, X = X, Y = Y)
  
  # The convergence value of each iteration calculated using the CostFuction
  CostFunctionValue <- CostFunction(parameters, X, Y)
  
  # Now I add the values to the dataframe 
  iterations[i,] <- c(i, CostFunctionValue)
}
```

```{r PLOT, message=FALSE}
ggplot(iterations,aes(x = iterations$Iterations , y = iterations$CostFunctionValue)) + geom_line() + xlab('Iterations') + ylab('CostFunctionValue')
```

# EXERCISE 3

I will look for some new ways of estimating the Gradient Descent using another methods of optim function. In order to get that, I will make a new function with 4 variables, this function is basically the TestGradientFunction but the fourth variable will be the optim method.

```{r METHODS OPTIM}
DifferentMethodsOptim <- function(iterations, X, Y, method){
  # Initialize (b, W)
  parameters <- rep(0, ncol(X))
  # Check evolution
  print(paste("Initial Cost Function value: ", 
              convergence <- c(CostFunction(parameters, X, Y)), sep = ""))
  
  # updating (b, W) using gradient update
  
  # Derive theta using gradient descent using optim function
  # Look for information about the "optim" function (there are other options)
  parameters_optimization <- optim(par = parameters, fn = CostFunction, X = X, Y = Y, 
                                   control = list(maxit = iterations), method = method)
  #set parameters
  parameters <- parameters_optimization$par
  
  # Check evolution
  print(paste("Final Cost Function value: ", 
              convergence <- c(CostFunction(parameters, X, Y)), sep = ""))

 return(parameters)
}
```

Now I´m going to try this new function.

```{r TRYING BFGS}
parameters2 <- DifferentMethodsOptim(1000, X, Y, "BFGS")
parameters2
```

```{r TRYING CG}
parameters3 <- DifferentMethodsOptim(1000, X, Y, "CG")
parameters3
```

```{r TRYING SANN}
parameters4 <- DifferentMethodsOptim(1000, X, Y, "SANN")
parameters4
```

Method Brent does not run because is only available for one dimensional optimization.

L-BFGS-B gives us an error related with the values of fn.
